---
title: "STA 138 Project Problem 2"
author: "Dandi Peng, Yuhan Ning"
date: "2/1/2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

## 1. Introduction
<br />
We are interested in how gender and death type are related in dataset 'horror'. In this project, we aim to check dependence of two variables with assumption that all samples are random.

## 2. Summary
```{r, echo = FALSE, out.width = "50%"}
horror = read.csv('horror.csv')
n_sum = nrow(horror)
table_data = table(horror)
y1 = table_data[1,1]
y2 = table_data[1,2]
y3 = table_data[1,3]
y4 = table_data[1,4]
n1 = sum(table_data[,1])
n2 = sum(table_data[,2])
n3 = sum(table_data[,3])
n4 = sum(table_data[,4])
barplot(table_data, beside = TRUE, main = 'Gender vs Death Type', legend.text = rownames(table_data), args.legend = list(x = 'topright')) 
table2 = table(horror$death, horror$gen)
mosaicplot(table2, main = 'Gender vs Death Type')
```
<br />
There are `r n_sum` samples in the dataset 'horror' and the two-way table of counts are shown below:
```{r, echo = FALSE}
table_data
```
<br />
We plot barplot and mosaicplot to catch a better glimpse of the counts in different categories. Clearly, for different death types, group female has highest number of deaths in level 'Other' and lowest number of deaths in 'Shot'. The uneven distribution of counts implies the two factors may not be independent. Therefore, we use hypothesis tests and confidence intervals to check further.

## 3. Analysis
**Pearson's Chi-Square Test**

$H0: P(female|BFT) = P(female|Other) = P(female|Shot) = P(female|Stabbed)$
<br />
$Ha:$ At least one probability in H0 is not the same.

Under the null hypothesis, $X^2$ follows chi-square distribution with d.f = (2-1)*(4-1) = 3
```{r, echo = FALSE}
pearson.test = chisq.test(table_data, correct = FALSE)
test.stat = pearson.test$statistic
p.val = pearson.test$p.value
expt = pearson.test$expected
stdres = pearson.test$stdres
x_sqr = (pearson.test$observed - pearson.test$expected)^2/pearson.test$expected
```
<br />
Test statistic is `r test.stat` and p-value is `r p.val`. 

We got table of expected value, table of standard residuals and table of $X^2$:
```{r, echo = FALSE}
expt
stdres
x_sqr
```
<br />
Stabbed differed the least from what was expected under the null. Shot tended to report less than expected under H0 and it contributed the most to the rejection of H0. 

**Wilson-Adjusted Bonferroni Corrected Confidence Intervals**

To know how gender depends on death types, we'd like to compare the probability of being female for different death types using Bonferroni corrected confidence interval. 

Since we have four groups (BFT, Other, Shot, Stabbed), we compare 6 differences in total (4 chooses 2 equals 6). 

$g = 6$; $alpha = 0.05$
```{r, echo = FALSE}
g = 6
alpha = 0.05
ci1 = prop.test(c(y1+1, y2+1), c(n1+2, n2+2), correct = FALSE, conf.level = 1 - alpha/g)$conf.int[1:2]
ci2 = prop.test(c(y1+1, y3+1), c(n1+2, n3+2), correct = FALSE, conf.level = 1 - alpha/g)$conf.int[1:2]
ci3 = prop.test(c(y1+1, y4+1), c(n1+2, n4+2), correct = FALSE, conf.level = 1 - alpha/g)$conf.int[1:2]
ci4 = prop.test(c(y2+1, y3+1), c(n2+2, n3+2), correct = FALSE, conf.level = 1 - alpha/g)$conf.int[1:2]
ci5 = prop.test(c(y2+1, y4+1), c(n2+2, n4+2), correct = FALSE, conf.level = 1 - alpha/g)$conf.int[1:2]
ci6 = prop.test(c(y3+1, y4+1), c(n3+2, n4+2), correct = FALSE, conf.level = 1 - alpha/g)$conf.int[1:2]
results = rbind(ci1, ci2, ci3, ci4, ci5, ci6)
colnames(results) = c('Lower','Upper')
rownames(results) = c('P(female|BFT)-P(female|Other)','P(female|BFT)-P(female|Shot)','P(female|BFT)-P(female|Stabbed)','P(female|Other)-P(female|Shot)','P(female|Other)-P(female|Stabbed)','P(female|Shot)-P(female|Stabbed)')
results
```

## 4. Interpretation
<br />
From Pearson's Test, p-value is `r p.val`. If gender and death type are independent, we would observe our data or more extreme with probability `r p.val`.
Since p value is much smaller than alpha, we reject H0 and conclude that gender and death types are dependent.

Continually, we compare 6 pairs of difference. 
Among the results:

1. Confidence Intervals for P(female|BFT)-P(female|Other), P(female|BFT)-P(female|Stabbed), P(female|Other)-P(female|Stabbed) contain 0;

2. P(female|BFT)-P(female|Shot), P(female|Other)-P(female|Shot) are larger than 0; 

3. P(female|Shot)-P(female|Stabbed) is smaller than 0.

We are overall 95% confident that the probabilities for female to die in BFT, Other and Stabbed are the same, which is larger than the probability of Shot. In other words, it is least probable for women to die of shot.

## 5. Conclusion
Based on above tests and calculations, we find that gender and death type are dependent. The probability for female to die of shot is less than other death types. The probability of other 3 death types for female is approximately the same.

### R Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```